% =====================================================================
% Section 1 — Introduction
% =====================================================================
\section{Introduction}\label{sec:intro}

Modern server-side web frameworks for the Java Virtual Machine (JVM)
have converged on a ``batteries-included'' model: a single dependency
pulls in an entire ecosystem of transitive libraries covering
persistence, serialisation, security, dependency injection, and more.
This model is undeniably productive, and frameworks such as Spring
Boot, Quarkus, and Micronaut are deployed at enormous scale. Yet the
convenience carries costs that are often invisible at design time and
expensive at operations time: a deployable artifact of tens of
megabytes, a transitive-dependency tree of dozens to hundreds of
libraries, a correspondingly large attack surface, a steep learning
curve, and non-trivial start-up latency.

These costs matter most precisely where resources are scarce or
scrutiny is high. In resource-constrained and edge deployments, image
size and memory footprint dominate density and cost. In regulated or
security-sensitive settings, every transitive dependency is a supply-chain
liability that must be audited and patched. In teaching, a framework
whose internals cannot be read in an afternoon obscures rather than
illuminates the mechanics of the web. Across all three, a smaller,
auditable, self-contained framework is not a nostalgic preference but a
practical requirement.

There is, however, a gap in the design space. Minimal ``micro''
frameworks reduce dependencies but typically delegate JSON, pooling,
scheduling, and security to third-party libraries, and treat security
as opt-in configuration layered on top of the request path. We are
aware of no mainstream full-stack MVC framework for Jakarta EE that is
simultaneously \emph{zero-dependency at runtime} and
\emph{secure-by-default at the pipeline level}.

This paper presents \textbf{JxMVC}, a full-stack MVC framework for
Jakarta~EE~10 that occupies this gap. JxMVC implements routing, data
access, connection pooling, JSON serialisation, caching, scheduling,
WebSocket support, metrics, validation, OpenAPI generation, rate
limiting, CSRF protection, password hashing, and an OAuth~2.0/OIDC
client \emph{entirely from scratch} on top of the JDK and the Jakarta
Servlet API. The compiled framework is a \SI{253}{\kilo\byte} JAR of
54 classes with zero third-party runtime dependencies (the only
additional production requirement is the JDBC driver of the database in
use). Security protections are applied by the request pipeline to
\emph{every} request rather than being opted into per route.

We do not claim novel algorithms or techniques. The contribution is a
\emph{design point}---radical minimalism, hardening-by-default, and
deliberate readability---together with an honest empirical evaluation
of what that design point costs and buys. Using a reproducible,
containerised, single-command benchmark harness under pinned-core
isolation on bare-metal Linux, we compare JxMVC against Spring Boot,
Quarkus, Micronaut, and Javalin, plus a GraalVM native-image build.
JxMVC produces the smallest JVM-mode deployable image of the group,
starts faster than Spring Boot and Micronaut, and sustains competitive
throughput with statistically indistinguishable tail latency, serving
\num{92759575} requests across the campaign with zero errors. Its one
clear trade-off is higher resident memory, a direct consequence of the
embedded servlet container.

\paragraph{Contributions.} This paper makes the following
contributions:
\begin{enumerate}
  \item \textbf{JxMVC}, a zero-dependency (\SI{253}{\kilo\byte})
    full-stack MVC framework for Jakarta~EE covering routing, DI,
    connection pooling, caching, scheduling, WebSocket, OpenAPI,
    validation, metrics, and OIDC (Section~\ref{sec:design}).
  \item A \textbf{security-by-default pipeline design}---protections
    applied by construction rather than by opt-in---and its
    library-free implementation (Sections~\ref{sec:design}
    and~\ref{sec:impl}).
  \item A \textbf{complete OAuth~2.0/OIDC client} (authorization code
    with PKCE) built using only the JDK (Section~\ref{sec:impl}).
  \item A \textbf{reproducible, honest empirical evaluation} against
    four mainstream frameworks and a GraalVM native build, packaged as
    a one-command Docker harness that is itself a reusable artifact
    (Section~\ref{sec:eval}).
  \item An \textbf{open-source artifact} verified by a 347-check,
    dependency-free test suite and released under the MIT license
    (Section~\ref{sec:conclusion}).
\end{enumerate}
