% =====================================================================
% Section 3 — Design and Architecture
% =====================================================================
\section{Design and Architecture}\label{sec:design}

\subsection{Principles}

Three principles govern the design of JxMVC and explain most of its
concrete decisions.

\begin{description}
  \item[Zero runtime dependencies.] Every capability is implemented on
    top of the JDK and the Jakarta Servlet API. No third-party library
    is required at runtime; the only additional production artifact is
    the JDBC driver of the target database. This bounds the artifact
    size (a \SI{253}{\kilo\byte} JAR of 54 classes) and eliminates the
    transitive-dependency supply-chain surface.
  \item[Security-by-default.] Protections are properties of the request
    pipeline, not of individual routes. A newly written controller is
    hardened before its author adds a single security annotation;
    relaxing a protection is explicit and local.
  \item[Explicit, readable API.] The framework favours direct,
    inspectable control flow over ``magic''. Data access reads columns
    directly from a result row rather than mapping to reflection-bound
    POJOs, and the entire core is small enough to be read and audited
    end to end---a property we treat as a first-class feature for
    teaching and for security review.
\end{description}

\subsection{The request pipeline}
\label{sec:pipeline}

The heart of JxMVC is a fixed, ordered request pipeline implemented in
the core servlet (\texttt{MainLxServlet}). Every request traverses the
same fifteen-stage sequence, shown in its real execution order in
Figure~\ref{fig:pipeline}, so cross-cutting concerns---sanitisation,
CORS, rate limiting, authentication, validation, compression,
metrics---are applied uniformly and by construction. Grouping by role,
the principal stages are:

\input{figures/pipeline}

\begin{enumerate}
  \item \textbf{Sanitiser} (\texttt{BaseSanitizer}) --- XSS-oriented
    cleaning of request parameters and arguments.
  \item \textbf{CORS resolver} (\texttt{BaseCorsResolver}) ---
    declarative, annotation-driven CORS policy.
  \item \textbf{Dispatcher} (\texttt{BaseDispatcher}) --- routing by
    annotations, path templates \texttt{\{var\}}, and the
    \texttt{/controller/action/args} convention, with correct
    \texttt{405} handling and \texttt{HEAD}$\rightarrow$\texttt{GET}.
  \item \textbf{Rate limiter} (\texttt{JxRateLimiter}) --- per-client
    rate limiting keyed by action.
  \item \textbf{Security} (\texttt{JxSecurity}) --- authentication and
    role checks via a pluggable \texttt{JxAuthProvider}, fail-closed
    when no provider is registered.
  \item \textbf{Filters} (\texttt{JxFilters}) --- global
    before/after hooks.
  \item \textbf{Controller} --- the developer's action, with access to
    data via \texttt{JxDB}, \texttt{JxRepository}, and
    \texttt{JxTransaction}.
  \item \textbf{Validation} (\texttt{JxValidation}) --- constraint
    checking of bound request bodies.
  \item \textbf{Result} (\texttt{ActionResult}) --- a
    \textsc{view}/\textsc{text}/\textsc{json}/\textsc{redirect}
    outcome.
  \item \textbf{Compression} (\texttt{JxGzip}) --- gzip with a
    threshold-based passthrough.
  \item \textbf{Metrics} (\texttt{JxMetrics}) --- per-route counters
    and timing.
\end{enumerate}

Because the security-relevant stages (sanitisation, rate limiting,
authentication, validation, body caps) precede and enclose the
developer's controller, the ``secure-by-default'' property is
structural rather than advisory.

\subsection{Subsystems}

\input{figures/architecture}

Figure~\ref{fig:architecture} shows how JxMVC layers its subsystems on
the JDK and the Jakarta Servlet API. All subsystems are built
in-framework. Data access (\texttt{JxDB})
issues JDBC directly with named parameters and returns row objects
whose columns are read explicitly, avoiding reflection-based mapping.
\texttt{JxPool} is a custom connection pool with configurable
revalidation; \texttt{JxRepository} adds generic CRUD with soft delete
and pagination; \texttt{JxTransaction} provides thread-local JDBC
transactions. \texttt{JxJson} is a from-scratch JSON parser and
serialiser with input limits, complemented by a variadic JSON builder
(\texttt{GenApi}). \texttt{JxCache} is an in-memory cache with TTL and
an entry cap; \texttt{JxScheduler} runs cron, fixed-rate, fixed-delay,
and one-shot tasks; \texttt{JxEventBus} is a synchronous event bus.
\texttt{JxMetrics} exposes per-route statistics, \texttt{JxOpenApi}
generates an OpenAPI~3.0 specification automatically, and
\texttt{JxWebSocket} provides WebSocket endpoints with rooms and
broadcast. On Java~21 and later, the framework detects virtual threads
for asynchronous work.

\subsection{Security by default}

Beyond the pipeline ordering, v3.4.0 ships a set of hardening measures
that are on (or fail-closed) by default and tunable through
configuration properties: security headers, fail-closed authentication
in the absence of a provider, per-action rate limiting (so an attacker
cannot evade limits by rotating a path variable), session-token CSRF
protection, request-body and JSON size caps that return \texttt{413}
when exceeded, anti open-redirect enforcement, HTML output encoding for
XSS defence at render time, and PBKDF2-SHA256 password hashing with
constant-time verification. Internal introspection endpoints
(\texttt{/jx/info}, \texttt{/jx/metrics}, \texttt{/jx/openapi}) return
\texttt{404} under the production profile unless explicitly exposed,
and the health endpoint omits sensitive fields. The implementation of
the most security-critical of these---rate limiting and the OIDC
client---is described in Section~\ref{sec:impl}.
